---
title: "Honor Values as Identity Content"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos=list(CRAN='http://cran.rstudio.com/'))
```

## 0. Overview

The study uses a Random Intercept Cross-Lagged Panel model to examine the relationship between identification with a regional group membership in southern Italy and honor-related values of reputation. 

Participants were 1,173 individuals (Mage = 16.70, SDage = 1.10; 43.8% female, 54.8% male, 1.4% did not specify their gender) from Campania, a southern region in Italy. Questionnaires were in Italian and items were measured on a 7-point Likert scale (1 = completely disagree, 7 = completely agree). 

## 0.1 Codebook

CNOTMIS: Participants could be matched across waves.
r-r3: average of identity items in waves 1, 2 and 3
h-h3: average of honor items in waves 1, 2 and 3
age: participants' age
gender: participants' gender (1 = Male, 2 = Female)

RegId1: I am pleased to think of myself as Campano
RegId2: I am proud I am Campano
RegId3: I Identify with other people who live in Campania
RegId4: I feel a sense of belonging to Campania

HonCon1:it is important to me that others see me as someone who deserves respect
HonCon2:it is important to me that others regard me as someone who is not to be disrespected
HonCon3:how others think of my family is important to me
HonCon4:caring about the implications of my actions for my family’s social image is important to me
HonCon5:it is important to me to defend my family from criticism

(Prefix T2-T3 indicate the same measure at the second and third wave respectively)

## 1. Install and load relevant packages 

```{r}
#install.packages("tidyverse")
#install.packages("lavaan")
#install.packages("semTools")
#install.packages("psych")
#install.packages("MBESS")

library(tidyverse)
library(lavaan)
library(semTools)
library(psych)
library(MBESS)

```

## 2. Load data and descriptive statistics
```{r}
#load("df.Rda")

psych::describe(df)
```

###2.1 Center Age and Gender

```{r}
df <- df %>% ungroup() %>% mutate(age.uncentered = age)

df <- df %>% mutate(age = age - mean(age, na.rm = TRUE))
```


```{r}
df<-df %>% mutate(gender.original = gender)
df<-df %>% mutate(gender = if_else(gender == 1, 0, 1))
```


## 3. Reliability

### 3.1 Regional identity

```{r}
# T1
Omega.RegIdT1 <-df[, c(18, 19, 20, 21)]
ci.reliability(data=Omega.RegIdT1, N = 1172, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)

# T2
Omega.RegIdT2 <-df[, c(14, 15, 16, 17)]
ci.reliability(data=Omega.RegIdT2, N = 1087, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)

# T3
Omega.RegIdT3 <-df[, c(10, 11, 12, 13)]
ci.reliability(data=Omega.RegIdT3, N = 1087, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)
```

## 3.2 Honor

```{r}
# T1
Omega.HonorT1 <-df[, c(32, 33, 34, 35, 36)]
ci.reliability(data=Omega.HonorT1, N = 1172, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)

# T2
Omega.HonorT2 <-df[, c(27, 28, 29, 30, 31)]
ci.reliability(data=Omega.HonorT2, N = 1087, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)

# T3
Omega.HonorT3 <-df[, c(22, 23, 24, 25, 26)]
ci.reliability(data=Omega.HonorT3, N = 1086, type = "omega", interval.type = "ml", B = 10000, conf.level = 0.95)
```

## 4. Test of longitudinal measurement invariance

### 4.1 Regional identity

####4.1.1 Configural model

```{r}
RegId.Conf <- 
  '
  RegIdT1 =~ RegId1 + RegId2 + RegId3 + RegId4
  RegIdT2 =~ T2RegId1 + T2RegId2 + T2RegId3 + T2RegId4
	RegIdT3 =~ T3RegId1 + T3RegId2 + T3RegId3 + T3RegId4
	
	#Method-specific factors
	
	M1 =~ 1*RegId2 + 1*T2RegId2 + 1*T3RegId2
	M2 =~ 1*RegId3 + 1*T2RegId3 + 1*T3RegId3
	M3 =~ 1*RegId4 + 1*T2RegId4 + 1*T3RegId4
	
	M1 + M2 + M3 ~~ 0*RegIdT1 + 0*RegIdT2 + 0*RegIdT3
	M1 ~~ 0*M2 + 0*M3
	M2 ~~ 0*M3
'

Fit.RegId.Conf <- cfa(RegId.Conf, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.RegId.Conf, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```


#### 4.1.2 Metric model

```{r}
RegId.Metr <- 
  '
  RegIdT1 =~ a*RegId1 + b*RegId2 + c*RegId3 + d*RegId4
  RegIdT2 =~ a*T2RegId1 + b*T2RegId2 + c*T2RegId3 + d*T2RegId4
	RegIdT3 =~ a*T3RegId1 + b*T3RegId2 + c*T3RegId3 + d*T3RegId4

#Method-specific factors
	
	M1 =~ 1*RegId2 + 1*T2RegId2 + 1*T3RegId2
	M2 =~ 1*RegId3 + 1*T2RegId3 + 1*T3RegId3
	M3 =~ 1*RegId4 + 1*T2RegId4 + 1*T3RegId4
	
	M1 + M2 + M3 ~~ 0*RegIdT1 + 0*RegIdT2 + 0*RegIdT3
	M1 ~~ 0*M2 + 0*M3
	M2 ~~ 0*M3
'

Fit.RegId.Metr <- cfa(RegId.Metr, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.RegId.Metr, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```


#### 4.1.3 Scalar model 

```{r}
RegId.Scal <- 
  '
  RegIdT1 =~ a*RegId1 + b*RegId2 + c*RegId3 + d*RegId4
  RegIdT2 =~ a*T2RegId1 + b*T2RegId2 + c*T2RegId3 + d*T2RegId4
	RegIdT3 =~ a*T3RegId1 + b*T3RegId2 + c*T3RegId3 + d*T3RegId4
	
	RegId1 + T2RegId1 + T3RegId1 ~ 0*1
	RegId2 + T2RegId2 + T3RegId2 ~ f*1
	RegId3 + T2RegId3 + T3RegId3 ~ g*1
	RegId4 + T2RegId4 + T3RegId4 ~ h*1
	
	RegIdT1 + RegIdT2 + RegIdT3 ~ 1

#Method-specific factors
	
	M1 =~ 1*RegId2 + 1*T2RegId2 + 1*T3RegId2
	M2 =~ 1*RegId3 + 1*T2RegId3 + 1*T3RegId3
	M3 =~ 1*RegId4 + 1*T2RegId4 + 1*T3RegId4
	
	M1 + M2 + M3 ~~ 0*RegIdT1 + 0*RegIdT2 + 0*RegIdT3
	M1 ~~ 0*M2 + 0*M3
	M2 ~~ 0*M3
'

Fit.RegId.Scal <- cfa(RegId.Scal, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.RegId.Scal, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)
```


####4.1.4 Model Comparison
```{r}
summary(compareFit(Fit.RegId.Conf,Fit.RegId.Metr, Fit.RegId.Scal))
```


### 4.2 Honor 

#### 4.2.1 Configural model

```{r}
Hon.Conf <- 
  '
  HonT1 =~ HonCon1 + HonCon2 + HonCon3 + HonCon4 + HonCon5 
  HonT2 =~ T2HonCon1 + T2HonCon2 + T2HonCon3 + T2HonCon4 + T2HonCon5
  HonT3 =~ T3HonCon1 + T3HonCon2 + T3HonCon3 + T3HonCon4 + T3HonCon5
	
	#Method-specific factors
	
	M1 =~ 1*HonCon2 + 1*T2HonCon2 + 1*T3HonCon2
	M2 =~ 1*HonCon3 + 1*T2HonCon3 + 1*T3HonCon3
	M3 =~ 1*HonCon4 + 1*T2HonCon4 + 1*T3HonCon4
	M4 =~ 1*HonCon5 + 1*T2HonCon5 + 1*T3HonCon5
	
	M1 + M2 + M3 + M4 ~~ 0*HonT1 + 0*HonT2 + 0*HonT3
	#M1 ~~ 0*M2 + 0*M3 + 0*M4
	#M2 ~~ 0*M3 + 0*M4
	#M3 ~~ 0*M4
'

Fit.Hon.Conf <- cfa(Hon.Conf, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.Hon.Conf, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```


#### 4.2.2 Metric model

```{r}
Hon.Metr <- 
  '
  HonT1 =~ a*HonCon1 + b*HonCon2 + c*HonCon3 + d*HonCon4 + e*HonCon5 
  HonT2 =~ a*T2HonCon1 + b*T2HonCon2 + c*T2HonCon3 + d*T2HonCon4 + e*T2HonCon5
  HonT3 =~ a*T3HonCon1 + b*T3HonCon2 + c*T3HonCon3 + d*T3HonCon4 + e*T3HonCon5
	
	#Method-specific factors
	
	M1 =~ 1*HonCon2 + 1*T2HonCon2 + 1*T3HonCon2
	M2 =~ 1*HonCon3 + 1*T2HonCon3 + 1*T3HonCon3
	M3 =~ 1*HonCon4 + 1*T2HonCon4 + 1*T3HonCon4
	M4 =~ 1*HonCon5 + 1*T2HonCon5 + 1*T3HonCon5
	
	M1 + M2 + M3 + M4 ~~ 0*HonT1 + 0*HonT2 + 0*HonT3
	#M1 ~~ 0*M2 + 0*M3 + 0*M4
	#M2 ~~ 0*M3 + 0*M4
	#M3 ~~ 0*M4
'

Fit.Hon.Metr <- cfa(Hon.Metr, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.Hon.Metr, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```

#### 4.2.3 Scalar model

```{r}
Hon.Scal <- 
  '
  HonT1 =~ a*HonCon1 + b*HonCon2 + c*HonCon3 + d*HonCon4 + e*HonCon5 
  HonT2 =~ a*T2HonCon1 + b*T2HonCon2 + c*T2HonCon3 + d*T2HonCon4 + e*T2HonCon5
  HonT3 =~ a*T3HonCon1 + b*T3HonCon2 + c*T3HonCon3 + d*T3HonCon4 + e*T3HonCon5
  
  HonCon1 + T2HonCon1 + T3HonCon1 ~ 0*1
  HonCon2 + T2HonCon2 + T3HonCon2 ~ f*1
  HonCon3 + T2HonCon3 + T3HonCon3 ~ g*1
  HonCon4 + T2HonCon4 + T3HonCon4 ~ h*1
  HonCon5 + T2HonCon5 + T3HonCon5 ~ i*1
  
  HonT1 + HonT2 + HonT3 ~ 1
	
	#Method-specific factors
	
	M1 =~ 1*HonCon2 + 1*T2HonCon2 + 1*T3HonCon2
	M2 =~ 1*HonCon3 + 1*T2HonCon3 + 1*T3HonCon3
	M3 =~ 1*HonCon4 + 1*T2HonCon4 + 1*T3HonCon4
	M4 =~ 1*HonCon5 + 1*T2HonCon5 + 1*T3HonCon5
	
	M1 + M2 + M3 + M4 ~~ 0*HonT1 + 0*HonT2 + 0*HonT3
	#M1 ~~ 0*M2 + 0*M3 + 0*M4
	#M2 ~~ 0*M3 + 0*M4
	#M3 ~~ 0*M4
'

Fit.Hon.Scal <- cfa(Hon.Scal, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(Fit.Hon.Scal, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

modindices(Fit.Hon.Scal, sort = TRUE, minimum.value = 4)

```

#### 4.2.4 Model comparison

```{r}
summary(compareFit(Fit.Hon.Conf, Fit.Hon.Metr, Fit.Hon.Scal))
```

## 5. Cross-lagged panel model (CLPM) to test for stationarity

### 5.1 Freely estimated CLPM 

```{r}

CLPM.free <- '
r2 ~ r + h
h2 ~ r + h

r3 ~ r2 + h2
h3 ~ r2 + h2

r ~~ h
r2 ~~ h2
r3 ~~ h3

r ~~ r
h ~~ h

r2 ~~ r2
h2 ~~ h2

r3 ~~ r3
h3 ~~ h3
'
CLPM.fit.free <- lavaan(CLPM.free, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(CLPM.fit.free, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```

### 5.2 CLPM with stability stationarity 

```{r}
CLPM.StabStat <- '
r2 ~ a*r + h
h2 ~ r + b*h

r3 ~ a*r2 + h2
h3 ~ r2 + b*h2

r ~~ h
r2 ~~ h2
r3 ~~ h3

r ~~ r
h ~~ h

r2 ~~ r2
h2 ~~ h2

r3 ~~ r3
h3 ~~ h3
'
CLPM.fit.StabStat <- lavaan(CLPM.StabStat, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(CLPM.fit.StabStat, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```

### 5.3 CLPM with full stationarity 

```{r}
CLPM.FullStat<- '
r2 ~ a*r + c*h
h2 ~ d*r + b*h
r3 ~ a*r2 + c*h2
h3 ~ d*r2 + b*h2

r ~~ h
r2 ~~ h2
r3 ~~ h3

r ~~ r
h ~~ h

r2 ~~ r2
h2 ~~ h2

r3 ~~ r3
h3 ~~ h3
'
CLPM.fit.FullStat <- lavaan(CLPM.FullStat, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(CLPM.fit.FullStat, standardized = TRUE, rsq = TRUE, fit.measures = TRUE, nd=3)

```

### 5.4 Model Comparison

```{r}
lavTestLRT(CLPM.fit.free, CLPM.fit.StabStat, CLPM.fit.FullStat)
```
Chi-square difference tests non-significant (i.e., model fit is not significantly worse for the more restricted model) between freely estimated model, stability stationarity model, and full stationarity model --> most restrictive model with full stability is assumed 

## 6. Random-intercept cross-lagged panel model (RI-CLPM)

### 6.1 RI-CLPM with full stationarity

#### 6.1.1 No control variables

```{r}
RICLPM <- 
'
#between paths

Br =~ 1*r + 1*r2 + 1*r3
Bh =~ 1*h + 1*h2 + 1*h3

#within paths

wr1 =~ 1*r
wr2 =~ 1*r2
wr3 =~ 1*r3

wh1 =~ 1*h
wh2 =~ 1*h2
wh3 =~ 1*h3

# set residual variance zero

r ~~ 0*r
r2 ~~ 0*r2
r3 ~~ 0*r3

h ~~ 0*h
h2 ~~ 0*h2
h3 ~~ 0*h3

#lagged effects within-person

wr2 ~ a*wr1 + c*wh1 
wh2 ~ d*wr1 + b*wh1
wr3 ~ a*wr2 + c*wh2
wh3 ~ d*wr2 + b*wh2

#covariance within-person 
wr1 ~~ wh1
wr2 ~~ wh2
wr3 ~~ wh3

#variance and covariance of random intercepts
Br ~~ Br
Bh ~~ Bh
Br ~~ Bh

# Residual variance of the within-person variance
wr1 ~~ wr1
wh1 ~~ wh1
wr2 ~~ wr2
wh2 ~~ wh2
wr3 ~~ wr3
wh3 ~~ wh3
'

RICLPM.fit <- lavaan(RICLPM, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(RICLPM.fit, 
        standardized = T,
        fit.measures = TRUE, 
        rsquare= TRUE,
        nd=3)

```

#### 6.1.2 with control variables
```{r}
RICLPM.r <- 
'
#between paths

Br =~ 1*r + 1*r2 + 1*r3
Bh =~ 1*h + 1*h2 + 1*h3

#within paths

wr1 =~ 1*r
wr2 =~ 1*r2
wr3 =~ 1*r3

wh1 =~ 1*h
wh2 =~ 1*h2
wh3 =~ 1*h3

# set residual variance zero

r ~~ 0*r
r2 ~~ 0*r2
r3 ~~ 0*r3

h ~~ 0*h
h2 ~~ 0*h2
h3 ~~ 0*h3

#lagged effects within-person

wr2 ~ a*wr1 + c*wh1 
wh2 ~ d*wr1 + b*wh1
wr3 ~ a*wr2 + c*wh2
wh3 ~ d*wr2 + b*wh2

#covariates
r + r2 + r3 + h + h2 + h3 ~ gender + age

#covariance within-person 
wr1 ~~ wh1
wr2 ~~ wh2
wr3 ~~ wh3

#variance and covariance of random intercepts
Br ~~ Br
Bh ~~ Bh
Br ~~ Bh

# Residual variance of the within-person variance
wr1 ~~ wr1
wh1 ~~ wh1
wr2 ~~ wr2
wh2 ~~ wh2
wr3 ~~ wr3
wh3 ~~ wh3
'

RICLPM.fit.r <- lavaan(RICLPM.r, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary(RICLPM.fit.r, 
        standardized = T,
        fit.measures = TRUE, 
        rsquare= TRUE,
        nd=3)

```

#### 6.1.3 Confidence intervals
```{r}
parameterEstimates(RICLPM.fit.r)
```

#### 6.1.4 Model comparison with CLPM

This relates to the CLPM and RI-CLPM without covariates
```{r}
lavTestLRT(CLPM.fit.FullStat, RICLPM.fit)
```


### 6.2 RI-CLPM with removed cross-lagged effect honor -> regional identity

```{r}
RICLPM.r_2 <- '
#between paths

Br =~ 1*r + 1*r2 + 1*r3
Bh =~ 1*h + 1*h2 + 1*h3

#within paths

wr1 =~ 1*r
wr2 =~ 1*r2
wr3 =~ 1*r3

wh1 =~ 1*h
wh2 =~ 1*h2
wh3 =~ 1*h3

# set residual variance zero

r ~~ 0*r
r2 ~~ 0*r2
r3 ~~ 0*r3

h ~~ 0*h
h2 ~~ 0*h2
h3 ~~ 0*h3

#lagged effects within-persor variance

wr2 ~ a*wr1 
wh2 ~ d*wr1 + b*wh1
wr3 ~ a*wr2
wh3 ~ d*wr2 + b*wh2

#wr2 ~ a*wr1 + c*wh1 
#wh2 ~ b*wh1
#wr3 ~ a*wr2 + c*wh2
#wh3 ~ b*wh2

#covariates
r + r2 + r3 + h + h2 + h3 ~ gender + age

#covariance within-person variance (first wave is covariance, secord and third wave are covariances of residuals)
wr1 ~~ wh1
wr2 ~~ wh2
wr3 ~~ wh3

#variance and covariance of random intercepts
Br ~~ Br
Bh ~~ Bh
Br ~~ Bh

# Residual variance of the within-person centred variance
wr1 ~~ wr1
wh1 ~~ wh1
wr2 ~~ wr2
wh2 ~~ wh2
wr3 ~~ wr3
wh3 ~~ wh3

'

RICLPM.fit.r_2 <- lavaan(RICLPM.r_2, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary (RICLPM.fit.r_2, 
        standardized = T,
        fit.measures = TRUE, 
        rsquare= TRUE,
        nd=3)
```

#### 6.2.1 Model comparison

```{r}
#Compare models

lavTestLRT(RICLPM.fit.r, RICLPM.fit.r_2)
```
Chi-square test nonsignificant - honor->identity cross-lagged path can be removed without significant deterioration of model fit

### 6.3 RI-CLPM with removed cross-lagged effect regional identity -> honor 

```{r}

RICLPM.r_3 <- '
#between paths

Br =~ 1*r + 1*r2 + 1*r3
Bh =~ 1*h + 1*h2 + 1*h3

#within paths

wr1 =~ 1*r
wr2 =~ 1*r2
wr3 =~ 1*r3

wh1 =~ 1*h
wh2 =~ 1*h2
wh3 =~ 1*h3

# set residual variance zero

r ~~ 0*r
r2 ~~ 0*r2
r3 ~~ 0*r3

h ~~ 0*h
h2 ~~ 0*h2
h3 ~~ 0*h3


#lagged effects within-person variance

wr2 ~ a*wr1 + c*wh1 
wh2 ~ b*wh1
wr3 ~ a*wr2 + c*wh2
wh3 ~ b*wh2

#covariates
r + r2 + r3 + h + h2 + h3 ~ gender + age

#covariance within-person variance 
wr1 ~~ wh1
wr2 ~~ wh2
wr3 ~~ wh3

#variance and covariance of random intercepts
Br ~~ Br
Bh ~~ Bh
Br ~~ Bh

# Residual variance of the within-person centred variance
wr1 ~~ wr1
wh1 ~~ wh1
wr2 ~~ wr2
wh2 ~~ wh2
wr3 ~~ wr3
wh3 ~~ wh3

'
RICLPM.fit.r_3 <- lavaan(RICLPM.r_3, data = df, estimator = 'ML', missing = 'ml.x', meanstructure = T, int.ov.free = T)

summary (RICLPM.fit.r_3, 
        standardized = T,
        fit.measures = TRUE, 
        rsquare= TRUE,
        nd=3)
```

#### 6.3.1 Model comparison

```{r}
#Compare models

lavTestLRT(RICLPM.fit.r, RICLPM.fit.r_3)
```
Chi-square test significant - removal of identity->honor cross-lagged path leads to significant deterioration of model fit
